\tw: Building a Xanalogical System in Ten Days

### 2003-02-13 papers/hypertext2003/paper.tex
Added section on micropayments.

\section{Introduction}
Xanadu has been floating in the collective mind of the hypertext research community for nearly forty years.
Xanadu is a dream and an ideal.  
We know it would be great (or at least interesting) if a xanalogical system existed, but we are also resigned to the notion that such a system will probably never exist.
Is Xanadu impossible?

@@

As mentioned in the above quote, Xanadu's underlying ideas were first incubated by Nelson in the 1960s \cite{Nelson1965}.  
These ideas evolved somewhat over the next forty years, receiving one of their most verbose presentations in Nelson's book {\it Literary Machines}, which was first published in 1981 \cite{NelsonLiteraryMachines}.  
Nelson's more recent paper on the xanalogical model is up to date in terms of modern technological trends, but the core ideas have not changed \cite{Nelson1999}.  
The high-level description and interpretation of Xanadu given here is based on information culled from {\it Literary Machines}, Nelson's recent lectures, and personal conversations with Nelson.

@@

\subsubsection{Development}
The entire \tw \  system is implemented as a single CGI script written in Perl, though the script is broken up into eight modules for organization purposes.
Perl's built-in support for regular expressions was very useful for the rather complicated text processing done by \tw.
Many of these text processing operations could be expressed in succinct and elegant ways using Perl.
For example, highlighting possible misspellings in preview mode was a minor addition that required less than fifteen lines of code.

CGI programming is well-supported in Perl, and HTML made user interface layout and user action processing easy to implement.
Perl-based CGI programming seems to be the modern method of choice for rapidly developing complex, multi-user applications.
Other systems that have used this development method successfully are discussed in Section \ref{sec:RelatedWork}.

@@

\subsection{Xanalogical Improvements}

\subsection{Missing Features}

\section{Micropayment Issues}

\subsection{Balancing Risk}

@@

>>>>

\section{Introduction}
\label{sec:Introduction}
Xanadu has been floating in the collective mind of the hypertext research community for nearly forty years.
Xanadu is a dream and an ideal.  
We know it would be great (or at least interesting) if a xanalogical system existed, but we are also resigned to the notion that such a system will probably never exist.
Is Xanadu impossible?

@@

\label{sec:Xanadu}
As mentioned in the above quote, Xanadu's underlying ideas were first incubated by Nelson in the 1960s \cite{Nelson1965}.  
These ideas evolved somewhat over the next forty years, receiving one of their most verbose presentations in Nelson's book {\it Literary Machines}, which was first published in 1981 \cite{NelsonLiteraryMachines}.  
Nelson's more recent paper on the xanalogical model is up to date in terms of modern technological trends, but the core ideas have not changed \cite{Nelson1999}.  
The high-level description and interpretation of Xanadu given here is based on information culled from {\it Literary Machines}, Nelson's recent lectures, and personal conversations with Nelson.

@@

\subsubsection{Development}
The entire \tw \  system is implemented as a single CGI script written in Perl, though the script is broken up into eight modules for organization purposes.
Perl's built-in support for regular expressions was useful for the rather complicated text processing done by \tw.
Many of these text processing operations can be expressed in succinct and elegant ways using Perl.
For example, highlighting possible misspellings in preview mode was a minor addition that required less than fifteen lines of code.

User interfaces can be layed out easily with HTML, and CGI processing is well-supported by existing Perl modules.
The CGI model requires a stateless action-response programming style, which makes it easy to focus on one feature at a time during development.
Perl-based CGI programming seems to be the modern method of choice for rapidly developing complex, multi-user applications.
Other systems that have used this development method successfully are discussed in Section \ref{sec:RelatedWork}.

@@

%\subsection{Xanalogical Improvements}

%\subsection{Missing Features}

\section{Micropayment Issues}
With our micropyament system, we wanted to leverage PayPal's \cite{paypal} immense popularity and familiarity.
By handling real-world currency exclusively through PayPal, instead of through our own virtual money system, we can process payments in a secure and trustworthy fashion.

However, PayPal's currency model does not support the microdollar resolution that we need, so we cannot rely on this service alone for transfering tokens between users.
Even if PayPal did support high-resolution payments, it does not allow a third party to initiate transfers between account holders in a transparent and frictionless way (and for good reason, given the security issues).

To deal with these issues, we built an internal token balance system and used PayPal integration only for deposits and withdrawls of real-world currency.
For example, a deposit is by no means frictionless (logging in to PayPal is required), but after the deposit has been made, token payments are completely transparent as the user accesses \tw \ documents.
Similarly, withdrawing tokens in exchange for real-world currency is not frictionless: 
PayPal does not support an automated payment mechanism, so withdrawl requests are queued for manual processing by the \tw \ system administrator.
Whereas deposits take effect immediately (using PayPal's ``Instant Payment Notification'' service), withdrawn token refunds are delayed by the response of the human administrator.
From the administrator's point of view, the withdrawl queue can be used to increase security, since any suspicious activity can be investigated before real-world payments are made.

\subsection{Balancing Risk}
Given that micropayments in \tw \ are mandatory and automatic, a major barrier to system adoption might exist:  trying \tw \ would require the deposit of real-world currency.
Even the main page displayed after login is a first-class \tw \ document, so a new user with no tokens would not be able to read a description of \tw, let alone try any of its features.

One way of dealing with this adoption barrier would be to give each new user a small number of ``free'' tokens to start with (for example, at least enough tokens to read the main page).
This solution would allow new users to try the system with no monetary investment, but it carries an enormous risk for the \tw \ system administrators.
A small group of users could create many fake accounts and then use each account to read a large dummy document owned by one user.
This kind of scam could be used to consolodate the free starting tokens from many accounts into one account, and then a withdrawl could be made in exchange for a large amount of real-world currency.
Even if users are not trying to intentially scam the system, doling out free tokens poses considerable risk.
One user might write documents that are popular with new users (for example, a collection of how-to documents) and ammass a large number of free tokens over time---the withdrawl of these tokens could financially devastate the \tw \ administrators.

We have devised a risk-free mechanism that deals with the adoption barrier.  
Our approach uses two token types.
The first type of tokens, called {\it trial} tokens, have no real-world value.
Each new account starts out with a small number of trial tokens (in our live system, each account starts with 50,000 trial tokens).
The second type of tokens, called {\it real} tokens, can be exchanged for real-world currency.
Token transfers inside \tw \ keeps track of token types:  trial tokens can only be transfered between users' trial balances, and real tokens can only be transfered between real balances.
Trial tokens are favored, and real tokens are only spent when a user has no trial tokens.

For example, suppose the balances of three users are as shown in following table:
%suppose user $A$ has a trial:real token balance of 50:200, user $B$ has a balance of 0:500, and user $C$ has a balance of 10:0.
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&50 &200\\
\hline
$B$&0  &500\\
\hline
$C$&10 &0  \\
\hline
\end{tabular}
\end{center}
Suppose that user $A$ reads a new 100-chracter document that is owned by $B$, but that the document contains a 20-character quote by $C$.
The resulting balances are shown in the following table:
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
User&Trial balance &Real balance\\
\hline
$A$&0 &150\\
\hline
$B$&50  &520\\
\hline
$C$&10 &20  \\
\hline
\end{tabular}
\end{center}

%$A$'s new balance would be 0:150, $B$'s new balance would be 50:520, and $C$'s new balance would be 10:20.

This two-type scheme allows users to try \tw \ extensively without a monetary investment, and it also poses no financial risks to the administrators.
Note that though trial tokens have no real-world value, they still have value inside \tw, since they can always be used to purchase new content.
Inside \tw, a user that amasses many trial tokens is just as wealthy as a user with many real tokens.
A user that actively published popular content in \tw \ could effectively use the system forever without depositing any real-world money.

@@

Much of the existing work on xanalogical systems was discussed in Sections \ref{sec:Introduction} and \ref{sec:Xanadu}.
